---
title: "Exercise #3"
subtitle: "Fortgeschrittene Statistische Software für NF - SS24"
author: "Serra Aksoy (12872071)"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exercise 1: Initializing git (4 Points)

For this whole exercise sheet we will be tracking all our changes to it
in git.

a)  Start by initializing a new R project with git support, called
    `2024-exeRcise-sheet-3`. If you forgot how to do this, you can
    follow this
    [guide](https://malikaihle.github.io/Introduction-RStudio-Git-GitHub/rstudio_project.html).
b)  Commit the files generated by Rstudio.
c)  For all of the following tasks in this exercise sheet we ask you to
    always commit your changes after finishing each subtask e.g. create
    a commit after task *1d*, *1e* etc.

> Note: This applies only to answers that have text or code as their
> answer. If you complete tasks in a different order or forget to commit
> one, this is no problem. If you change your answers you can just
> create multiple commits to track the changes.

d)  Name 2 strengths and 2 weaknesses of git. (Don't forget to create a
    commit after this answer, see *1c*)
    
**Strengths:**
**Distributed Version Control:** Git is a distributed version control system, which means every developer has a full copy of the repository, including its history. This makes it resilient to network failures and allows for more flexibility in workflows, such as offline work and easy branching.

**Branching and Merging:** Git provides robust support for branching and merging. Developers can create branches to work on new features or fixes independently without affecting the main codebase. Git's merging capabilities allow for easy integration of changes from one branch to another, facilitating collaboration among team members.

**Weaknesses:**
**Learning Curve:** Git has a steep learning curve, especially for beginners. Its powerful feature set and command-line interface can be overwhelming at first. Understanding concepts like branching, merging, rebasing, and resolving conflicts requires time and practice.

**Complexity in Resolving Conflicts:** While Git offers powerful conflict resolution tools, resolving conflicts can sometimes be complex, especially in scenarios involving multiple branches and conflicting changes. This complexity increases with the size and complexity of the project, requiring careful attention to ensure changes are integrated correctly.
    
e)  Knit this exercise sheet. Some new files will automatically be
    generated when knitting the sheet e.g. the HTML page. Ignore these
    files, as we only want to track the source files themselves.

## Exercise 2: Putting your Repository on GitHub (3.5 Points)

For this task you will upload your solution to GitHub.

a)  Create a new repository on GitHub in your account named
    `exeRcise-sheet-3`. Make sure you create a **public repository** so
    we are able to see it for grading. Add the link to the repository
    below:
    **Repository link:** "https://github.com/Serurays/exeRcise-sheet-3"
    
b)  Push your code to this new repository by copying and executing the
    snippet on github listed under
    `…or push an existing repository from the command line`.
c)  Regularly push your latest changes to GitHub again and especially do
    so when you are finished with this sheet.

## Exercise 3: Baby-Names in Munich (3.5 Points)

Download the latest open datasets on given names ("Vornamen") from the
open data repository of the city of Munich for the years `2023` and
`2022`.

Link: <https://opendata.muenchen.de/dataset/vornamen-von-neugeborenen>

a)  Download the data for both years and track it in git. For small
    datasets like these adding them to git is not a problem.

b)  Load the data for both years into R. Check the type of the count
    variable ("Anzahl") and look into the data to determine why it is
    not numeric? Fix the problem in an appropriate manner, it is OK if
    some of the counts are inaccurate because of this. Explain your
    solution and the repercussions.
    
```{r load_data}
  
data_2022 <- read.csv("data/vornamen_muenchen_2022.csv")
data_2023 <- read.csv("data/vornamen_muenchen_2023.csv")
```

```{r check_type_of_anzahl}
  
class(data_2022$Anzahl)
class(data_2023$Anzahl)
```

```{r determine_problem}

head(data_2022$Anzahl)
head(data_2023$Anzahl)
tail(data_2022$Anzahl)
tail(data_2023$Anzahl)
```
```{r solve_problem_by_turning_variable_to_numeric}

clean_anzahl <- function(data) {
  data$Anzahl <- as.numeric(gsub("[^0-9]", "", data$Anzahl))
  data$Anzahl[is.na(data$Anzahl)] <- 4
  return(data)
}

data_2022 <- clean_anzahl(data_2022)
data_2023 <- clean_anzahl(data_2023)

head(data_2022$Anzahl)
head(data_2023$Anzahl)
tail(data_2022$Anzahl)
tail(data_2023$Anzahl)
```

**Solution:** The function clean_anzahl uses gsub to remove any non-numeric characters from the "Anzahl" variable, leaving only the numeric values.
If the value was "4 oder weniger", it replaces the resulting NA with 4, assuming this means "4 or fewer". This approach ensures that the "Anzahl" variable contains numeric values, making it suitable for further analysis. However, it does make an assumption about the meaning of "4 oder weniger" and replaces it with 4. If this assumption is incorrect, it might introduce inaccuracies in the data.

c)  Calculate the total number of babies born in Munich in 2022
    and 2023. Which year had the bigger baby-boom?
    
```{r calculate_compare_num_babies}

total_2022 <- sum(data_2022$Anzahl)
total_2023 <- sum(data_2023$Anzahl)

if (total_2022 > total_2023) {
  bigger_boom <- "2022"
} else if (total_2023 > total_2022) {
  bigger_boom <- "2023"
} else {
  bigger_boom <- "Both years had the same number of babies"
}

cat("Total number of babies born in 2022:", total_2022, "\n")
cat("Total number of babies born in 2023:", total_2023, "\n")
cat("Year with the bigger baby-boom:", bigger_boom, "\n")
```

d)  Add a new column `year` to both datasets which holds the correct
    year for each.
    
```{r add_year_column, message=FALSE, warning=FALSE}

library(tidyverse)

data_2022 <- data_2022 %>%
  mutate(year = 2022)

data_2023 <- data_2023 %>%
  mutate(year = 2023)
```

e)  Combine both datasets into one using `bind_rows()`.

```{r combine_datasets}

combined_data <- bind_rows(data_2022, data_2023)
head(combined_data)
tail(combined_data)
```

f)  Combine the counts for same names to determine the most popular
    names across both years. Print out the top 10 names in a nicely
    formatted table for both years. Include a table caption.
    
```{r most_popular_names, message=FALSE, warning=FALSE}

library(kableExtra)

combined_names <- combined_data %>%
  group_by(Vorname) %>%
  summarise(Total_Count = sum(Anzahl)) %>%
  arrange(desc(Total_Count))

top_10_names <- combined_names %>% 
  head(10)

top_10_names %>%
  knitr::kable(caption="Top 10 Most Popular Baby Names in Munich for 2022 and 2023") %>%
  kable_styling(bootstrap_options = "striped",
                full_width = FALSE)
```

## Exercise 4: Open Analysis (4 points)

This exercise is a bit more open-ended. You can choose any dataset from
[Our World in Data](https://ourworldindata.org/) and analyze it, while
determining the research question yourself.

a)  Go to <https://github.com/owid/owid-datasets/tree/master/datasets>
    and choose a dataset that interests you. You can have a look at
    <https://ourworldindata.org/> to gather some inspiration.
    
b)  Download the dataset and track it in git.

```{r download_dataset, message=FALSE, warning=FALSE}

data <- read_csv("https://raw.githubusercontent.com/owid/owid-datasets/master/datasets/Antibiotic%20use%20in%20livestock%20-%20European%20Commission%20%26%20Van%20Boeckel%20et%20al./Antibiotic%20use%20in%20livestock%20-%20European%20Commission%20%26%20Van%20Boeckel%20et%20al..csv")

head(data)
```

c)  Put the name / title of the dataset and a link to it below.
-   **Dataset Name:** "Antibiotic use in livestock - European Commission & Van Boeckel et al."
-   **Link:** <https://github.com/owid/owid-datasets/tree/master/datasets/Antibiotic%20use%20in%20livestock%20-%20European%20Commission%20%26%20Van%20Boeckel%20et%20al.>

d)  Come up with a (research) question you want to answer with the data
    and briefly explain why you believe this is an interesting question
    within one sentence. It should be a question that can be answered
    with the dataset and using R.
    
**Research Question:** How has the antibiotic use in livestock changed over time in Germany from 2010 to 2015, and how does it compare to other countries in the dataset for the year 2010?

**Explanation:** This question is interesting because it can help to understand the trend of antibiotic use in livestock over time in Germany and compare it with other countries' usage in a specific year (2010). This can provide insights into the effectiveness of policies or practices aimed at reducing antibiotic use in livestock.

e)  Use R to answer your chosen question.

```{r slice_needed_data}

germany_data <- data %>% filter(Entity == "Germany")
data_2010 <- data %>% filter(Year == 2010)

colnames(germany_data) <- c("Entity", "Year", "Antibiotic_Use")
colnames(data_2010) <- c("Entity", "Year", "Antibiotic_Use")

head(germany_data)
head(data_2010)
```

**Answer for the first part of the Research Question:** From the table for Germany, we can see that the antibiotic use in livestock remained constant at 211.5 mg/kg in 2010 and 2011. It then started to decline from 204.8 mg/kg in 2012 to 97.9 mg/kg in 2015. This indicates a significant reduction in antibiotic use over the six-year period. This decline can be attributed to a series of targeted policies and initiatives, including the German Antibiotic Resistance Strategy (DART), the 16th Amendment of the German Medicinal Products Act, and the establishment of reduction targets for antibiotic use in agriculture. Enhanced monitoring through a national database, increased research efforts, and awareness campaigns further supported these regulatory measures.

**To address the second part of the question, it's beneficial to sort the data and see where Germany stands.**

```{r sort_data_2010}

sorted_data_2010 <- data_2010 %>% 
  arrange(desc(Antibiotic_Use))

germany_index <- which(sorted_data_2010$Entity == "Germany")

sorted_data_2010
germany_index
```

**Answer for the second part of the Research Question:** In 2010, Germany ranked 5th among all the countries in the dataset in terms of antibiotic use in livestock, with a usage rate of 211.5 mg/kg. This indicates that, at that time, Germany had one of the higher antibiotic usage rates compared to other countries included in the dataset. However, it's important to note that several other countries (Italy, Cyprus, Hungary, Spain) had even higher usage rates, suggesting that antibiotic use in livestock was a widespread issue across multiple regions. This comparison underscores the significance of monitoring and addressing antibiotic use practices in livestock farming on a global scale to mitigate the risk of antibiotic resistance and safeguard public health.

f)  Create a meaningful plot / figure with the dataset. Make sure to
    provide a figure caption (via the chunk options / Rmarkdown) and
    correctly label the figure.
    
```{r plot_antibiotic_use_germany, fig.cap="Figure 1: Antibiotic Use in Livestock in Germany (2010-2015)", warning=FALSE, message=FALSE}

library(ggplot2)

ggplot(germany_data, aes(x = Year, y = Antibiotic_Use)) +
  geom_line(color = "blue") +
  geom_point(color = "blue") +
  labs(x = "Year",
       y = "Antibiotic Use (mg/kg)") +
  theme_minimal()
```

```{r plot_antibiotic_use_2010, fig.cap="Figure 2: Top 10 Countries with Highest Antibiotic Use in Livestock (2010)"}

top_countries_2010 <- data_2010 %>%
  arrange(desc(Antibiotic_Use)) %>%
  head(10)

# Plot antibiotic use in top countries
ggplot(top_countries_2010, aes(x = reorder(Entity, Antibiotic_Use), y = Antibiotic_Use)) +
  geom_bar(stat = "identity", fill = "lightblue") +
  coord_flip() +
  labs(x = "Country",
       y = "Antibiotic Use (mg/kg)") +
  theme_minimal()
```